---
title: "Longitudinal data model"
author: "Biyi Shen"
date: "2019/7/23"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```
# preliminary for the data
we use the dataset from ds_cgm_complete_3days.rds, and pick 60 patients (ID 5, 7, 11, 18, 21, 24, 25, 26, 29, 32, 34, 35, 41, 46, 47, 49, 50, 53, 55, 69, 74, 82, 83, 88, 89, 91, 94, 96, 101, 118, 120, 124, 127, 129, 133, 135, 150, 153, 156, 158, 159, 164, 167, 168, 171, 176, 184, 190, 191, 192, 194, 203, 204, 205, 207, 212, 213, 219, 220, 225)
```{r}
data<-readRDS(file = "../data/ds_cgm_complete_3days.rds")
#choose 30 from each group
used.id<-c(5, 7, 11, 18, 21, 24, 25, 26, 29, 32, 34, 35, 41, 46, 47, 49, 50, 53, 55, 69, 74, 82, 83, 88, 89, 91, 94, 96, 101, 118, 120, 124, 127, 129, 133, 135, 150, 153, 156, 158, 159, 164, 167, 168, 171, 176, 184, 190, 191, 192, 194, 203, 204, 205, 207, 212, 213, 219, 220, 225)
data.subset<-as.data.frame(subset(data, data$pt_id %in% used.id))
data.subset<-subset(data.subset, data.subset$visitnum %in% c(1,2,3))
data.subset$day.in.visit<-data.subset$day_in_subrange
data.subset$group<-ifelse(data.subset$trt_group=="CGM Only",0, 1)
data.subset$device_tm_order_in_day<-data.subset$device_tm_bin
# 
```

#  normal GEE
we want to detect the group difference in general in fixed visit
```{r}
data.subset<-data
#summary(gee(glucose_value ~ group, family=gaussian,data=data.subset, id=pt_id, corstr = "independence"))$coefficients

result.gee1<-matrix(NA, nrow = 6,ncol = 4)
for (i in 1:6) {
  data.used<-subset(data.subset,data.subset$visitnum==i-1)
  gee1<-summary(gee(glucose_value ~ group, family=gaussian,data=data.used, id=pt_id, corstr = "independence"))$coefficients
  result.gee1[i,1]<-i-1
  result.gee1[i,2]<-gee1[2,1]
  result.gee1[i,3]<-gee1[2,1]-1.96*gee1[2,4]
  result.gee1[i,4]<-gee1[2,1]+1.96*gee1[2,4]
  print(i)
}

colnames(result.gee1)<-c("visit","effect","low.ci","high.ci")
result.gee1<-as.data.frame(result.gee1)
ggplot(result.gee1, aes(x=visit, y=effect, group=1)) + geom_hline(yintercept=0, linetype="dashed", color = "blue")+
  geom_point(size=10,alpha=0.52) +
  geom_errorbar(width=.1, aes(ymin=low.ci, ymax=high.ci), colour="darkred") + 
  labs(x="visit",y= "group effect difference", title="Group difference estimation and CI with GEE", caption="source: A Randomized Trial Comparing CGM With and Without Routine BGM in Adults with T1D.") 

```
# Time varying model with GEE
We need to construct the spline by choosing knots and degree, and treating the spline matrix as part of covariate matrix. So what we get from gee is the estimate of coefficients in basis. We need calculation in order to get the estimate for $\beta(t)$ and corresponding variance and covariance estimation matrix. 
```{r}

### in specific visit####

choose.visit<-3
data.used<-subset(data.subset,data.subset$visitnum==choose.visit)

t <-data.used$device_tm_order_in_day
knots.choose<- round(288*(c(1:6)/7))
bsMat <- bSpline(t, knots = knots.choose, degree = 3, intercept=T)
data.used$group_spline_day<-data.used$group*bsMat
data.used$bsMat<-bsMat
gee2<-gee(glucose_value ~ bsMat+group_spline_day-1, family=gaussian,data=data.used, id=pt_id,corstr="exchangeable")

spline_beta0<-matrix((summary(gee2)$coefficients)[c(1:10),1],nrow=10,ncol = 1)
spline_beta<-matrix((summary(gee2)$coefficients)[c(11:20),1],nrow=10,ncol = 1)
se_beta0<-gee2$naive.variance[c(1:10),c(1:10)]
se_beta1<-gee2$naive.variance[c(1:10),c(1:10)]+gee2$naive.variance[c(11:20),c(11:20)]+2*gee2$naive.variance[c(1:10),c(11:20)]

estimated_matrix0<-bsMat%*%spline_beta0
se_fitted0<-sqrt(diag(bsMat%*%se_beta0%*%t(bsMat)))

estimated_matrix<-bsMat%*%spline_beta

estimated_matrix1<-estimated_matrix0+estimated_matrix
se_fitted1<-sqrt(diag(bsMat%*%se_beta1%*%t(bsMat)))

uplimit0<-estimated_matrix0+se_fitted0*1.96
lowlimit0<-estimated_matrix0-se_fitted0*1.96

uplimit1<-estimated_matrix1+se_fitted1*1.96
lowlimit1<-estimated_matrix1-se_fitted1*1.96

plot.data.1<-as.data.frame(cbind(data.used$device_tm,estimated_matrix0,uplimit0,lowlimit0))
plot.data.2<-as.data.frame(cbind(data.used$device_tm,estimated_matrix1,uplimit1,lowlimit1))
plot.data.1$group.name<-"CGM"
plot.data.2$group.name<-"CGM+BGM"
plot.data<-rbind(plot.data.1,plot.data.2)

colnames(plot.data)<-c("t","fit","upper","lower","group.name")

plot.new<-plot.data %>%
  mutate(tclock = hms::as_hms(t)) 

 plot.new$tclock<-as.POSIXct(plot.new$tclock)

ggplot(plot.new, aes(x=tclock, y=fit,group = group.name,color=group.name))+ geom_line(aes(color=group.name))+
  geom_smooth(aes(ymin = lower, 
                  ymax = upper),
              stat = "identity") +
  scale_x_datetime(date_labels = "%H:%M", date_breaks = "6 hours") +
  xlab("time in mentoring(in visit 3)") +
  ylab("treatment effect")


```

# Time varying model with GAMM
Using package mgcv and gamm function.
```{r} 

data.used$group<-as.factor(data.used$group)
gammm1<-gamm(glucose_value ~ group+s(device_tm_order_in_day, bs= "cc", by=group,k=7) ,family=gaussian, data=data.used, correlation=corAR1(form=~1|visitnum), random = list(pt_id=~1,visitnum = ~1))
 
mm3 <- expand.grid(device_tm_order_in_day=seq(0,288,length=1000),group=levels(data.used$group))
pp3 <- predict.gam(gammm1$gam,mm3,se.fit=T,type='response')

rr3 <- data.frame(dv="2 exac2",mm3,pp3[1],pp3[2])
rr3$lower <- with(rr3,fit- qnorm(0.95)*se.fit)
rr3$upper <- with(rr3,fit+ qnorm(0.95)*se.fit)
rr3$group.name<-ifelse(rr3$group==0,"CGM","CGM+BGM") 

rr3$t<-rr3$device_tm_order_in_day*5*60
plot.new<-rr3 %>%
  mutate(tclock = hms::as_hms(t)) 

plot.new$tclock<-as.POSIXct(plot.new$tclock)

ggplot(plot.new, aes(x=tclock, y=fit,group = group.name,color=group.name))+ geom_line(aes(color=group.name))+
  geom_smooth(aes(ymin = lower, 
                  ymax = upper),
              stat = "identity") +
  scale_x_datetime(date_labels = "%H:%M", date_breaks = "6 hours") +
  xlab("time in mentoring(day)") +
  ylab("treatment effect")
 
```
