---
title: "Exploratory Work"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
---

```{r setup, include = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Loads libraries
library(tidyverse)
```

# Datasets

```{r}
# Loads in datasets
ds_cgm <- readRDS("../data/ds_cgm.RDS")
ds_subject <- readRDS("../data/ds_subject.RDS")
ds_visit <- readRDS("../data/ds_visit.RDS")
```

## ds_cgm

**Variables**

- pt_id: patient id
- visit: visit
- visitnum: visit number
- visit_dt_days_from_enroll_lag7: 
- visit_dt_days_from_enroll:
- device_dt_tm_days_from_enroll: 
- device_tm: time of day when observation is taken
- glucose_value: glucose value measured by device

```{r}
# Glimpses the ds_cgm dataset
glimpse(ds_cgm)
```

## ds_subject

**Variables** 

- pt_id: patient id
- site_id: 
- gender: gender of patient
- ethnicity: ethnicity of patient
- diag_age: age diagnosed with diabetes
- oth_gluc_lower_med: 
- edu_level: education level of patient
- weight: weight of patient (at start of study?)
- height: height of patient (at start of study?)
- cgm_use_status: information about whether the patient has used cgm before
- pt_status: 
- trt_group: treatment group (cgm or both)
- age_as_of_enroll_dt: age at enrollment date

```{r}
# Glimpses the ds_cgm dataset
glimpse(ds_subject)
```

## ds_visit

**Variables** 

- pt_id: patient id
- visit: 
- visitnum: 
- visit_dt_days_from_enroll: 
- visit_dt_days_from_enroll_lag7:

```{r}
# Glimpses the ds_cgm dataset
glimpse(ds_visit)
```


